---
title: "NCAA Bracket Prediction"
author: "Jack Crilly and Brian Papiernik"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries
```{r}
source("elo_funcs.r") # elo package
library(tidyverse)
library(stringr)
library(plyr)
library(hoopR)

library(gtools) # Load gtools package
library(RCurl) # Load RCurl
library(png) # Load jpeg
library(grid) # Load grid
library(ggplot2) # Load ggplot2
library(magick)
```
### Acquire Data

Acquire 2023-24 NCAA game team box score data
```{r}
temp <- load_mbb_team_box()

# gameid = id
temp <- temp %>%
  mutate(game_date = as.Date(game_date)) %>%
  filter(game_date <= "2024-03-17")

away_rows <- temp[which(temp$team_home_away == "away"), c(1,2,4,
                                                          6,9:11,18)]
home_rows <- temp[which(temp$team_home_away == "home"), c(1,2,4,
                                                          6,9:11,18)]

names(away_rows)[c(4:8)] <- c("away_id", "away_team", "away_mascot", "away_abbreviation", "away_score")
names(home_rows)[c(4:8)] <- c("home_id", "home_team", "home_mascot", "home_abbreviation", "home_score")

cbb_season_results <- cbind(home_rows, away_rows)

cbb_season_results <- cbb_season_results[,-c(9,10,11)]

cbb_season_results$result <- NA

cbb_season_results$result[which(cbb_season_results$home_score > cbb_season_results$away_score)] <- 1
cbb_season_results$result[which(cbb_season_results$home_score < cbb_season_results$away_score)] <- 0
```

# Elo Prediction

### Calculate ELO Ratings
```{r}
teams <- unique(c(cbb_season_results$home_team, cbb_season_results$away_team))
```

We then need to assign an initial Elo rating for each team, we will give each team an initial value of 1,500:

```{r}
# Join teams and initial elo rating
teams_elo <- cbind.data.frame(teams, rep(1500, length(teams)))
# Name columns
names(teams_elo) <- c("teams", "elo")
# View first few rows of teams
head(teams_elo)
```

We are now ready to assign calculate Elo ratings using the historical data. As our data is in chronological order we will work from the first row and update the Elo ratings one game at a time:

```{r}
# For each game in our dataset
for(i in 1:nrow(cbb_season_results)){
  # Extract match
  game <- cbb_season_results[i,]
  
  # Extract team 1 Elo
  team1_elo <- teams_elo$elo[teams_elo$team == game$home_team]
  # Extract team 2 Elo
  team2_elo <- teams_elo$elo[teams_elo$team == game$away_team]
  
  # Calculate new Elo ratings
  new_elo <- elo.calc(wins.A = game$result,
                      elo.A = team1_elo,
                      elo.B = team2_elo,
                      k= 50)
  
  # Store new elo ratings for home team
  teams_elo$elo[teams_elo$team == game$home_team] <- new_elo[1,1]
  # Store new elo ratings for away team
  teams_elo$elo[teams_elo$team == game$away_team] <- new_elo[1,2]
}
```

Top teams by ELO
```{r}
teams_elo <- teams_elo[order(teams_elo$elo, decreasing = TRUE),]
teams_elo
```

### March Madness Bracket

Create a vector of March Madness teams
```{r}
mm_teams <- c(#East:
  "UConn", # 1
  "Stetson", # 16
  "Iowa State", # 2
  "South Dakota State", # 15
  "Illinois", # 3 
  "Morehead State", # 14
  "Auburn", # 4
  "Yale", # 13
  "San Diego State", # 5
  "UAB", # 12
  "BYU", # 6
  "Duquesne", # 11
  "Washington State", # 7 
  "Drake", # 10
  "Florida Atlantic", # 8 
  "Northwestern", # 9  
  
  #West:
  "North Carolina", # 1
  "Wagner", # 16 
  "Arizona", # 2
  "Long Beach State", # 15
  "Baylor", # 3
  "Colgate", # 14
  "Alabama", # 4
  "Charleston", # 13
  "Saint Mary's", # 5
  "Grand Canyon", # 12
  "Clemson", # 6
  "New Mexico", # 11 
  "Dayton", # 7
  "Nevada", # 10
  "Mississippi State", # 8
  "Michigan State", # 9
  
  #South:
  "Houston", # 1
  "Longwood", # 16
  "Marquette", # 2
  "Western Kentucky", # 15
  "Kentucky", # 3
  "Oakland", # 14
  "Duke", # 4
  "Vermont", # 13
  "Wisconsin", # 5
  "James Madison", # 12
  "Texas Tech", # 6
  "NC State", # 11
  "Florida", # 7
  "Colorado", # 10, or Boise State
  "Nebraska", # 8
  "Texas A&M", # 9
  
  #Midwest:
  "Purdue", # 1
  "Grambling", # 16
  "Tennessee", # 2
  "Saint Peter's", # 15
  "Creighton", # 3
  "Akron", # 14
  "Kansas", # 4
  "Samford", # 13
  "Gonzaga", # 5
  "McNeese", # 12
  "South Carolina", # 6
  "Oregon", # 11
  "Texas", # 7
  "Colorado State",
  "Utah State", # 8
  "TCU" # 9
  )
```

Acquire Elo ratings for tournament teams
```{r}
mm_elo <- data.frame("teams" = mm_teams, "elo" = 1500)

for( i in 1:64){
  mm_elo$elo[i] <- teams_elo[teams_elo$teams == mm_elo$teams[i],]$elo
}
```

Create tournament system
```{r}
Tourney24 <-  data.frame(teams = c(#East:
  "UConn", # 1
  "Stetson", # 16
  "Iowa State", # 2
  "South Dakota State", # 15
  "Illinois", # 3 
  "Morehead State", # 14
  "Auburn", # 4
  "Yale", # 13
  "San Diego State", # 5
  "UAB", # 12
  "BYU", # 6
  "Duquesne", # 11
  "Washington State", # 7 
  "Drake", # 10
  "Florida Atlantic", # 8 
  "Northwestern", # 9  
  
  #West:
  "North Carolina", # 1
  "Wagner", # 16 
  "Arizona", # 2
  "Long Beach State", # 15
  "Baylor", # 3
  "Colgate", # 14
  "Alabama", # 4
  "Charleston", # 13
  "Saint Mary's", # 5
  "Grand Canyon", # 12
  "Clemson", # 6
  "New Mexico", # 11 
  "Dayton", # 7
  "Nevada", # 10
  "Mississippi State", # 8
  "Michigan State", # 9
  
  #South:
  "Houston", # 1
  "Longwood", # 16
  "Marquette", # 2
  "Western Kentucky", # 15
  "Kentucky", # 3
  "Oakland", # 14
  "Duke", # 4
  "Vermont", # 13
  "Wisconsin", # 5
  "James Madison", # 12
  "Texas Tech", # 6
  "NC State", # 11
  "Florida", # 7
  "Colorado", # 10, or Boise State
  "Nebraska", # 8
  "Texas A&M", # 9
  
  #Midwest:
  "Purdue", # 1
  "Grambling", # 16
  "Tennessee", # 2
  "Saint Peter's", # 15
  "Creighton", # 3
  "Akron", # 14
  "Kansas", # 4
  "Samford", # 13
  "Gonzaga", # 5
  "McNeese", # 12
  "South Carolina", # 6
  "Oregon", # 11
  "Texas", # 7
  "Colorado State", # 10
  "Utah State", # 8
  "TCU" # 9
  ), 
  
                       seed_s = c("W01", "W16", "W02", "W15", "W03", "W14", "W04", "W13",
                                  "W05", "W12", "W06", "W11", "W07", "W10", "W08", "W09",
                                  "X01", "X16", "X02", "X15", "X03", "X14", "X04", "X13",
                                  "X05", "X12", "X06", "X11", "X07", "X10", "X08", "X09",
                                  "Y01", "Y16", "Y02", "Y15", "Y03", "Y14", "Y04", "Y13", 
                                  "Y05", "Y12", "Y06", "Y11", "Y07", "Y10", "Y08", "Y09",
                                  "Z01", "Z16", "Z02", "Z15", "Z03", "Z14", "Z04", "Z13",
                                  "Z05", "Z12", "Z06", "Z11", "Z07", "Z10", "Z08", "Z09"
                                  ), 
                       
                       
                       # Work from outsides of range building into the median to create WO1 vs. WO16 and so on reference the march madness bracket. 
                       
                       Game1 = c( 1,1,   2,2,   3,3,   4,4,   5,5,   6,6,   7,7,   8,8,
                                  9,9,  10,10, 11,11, 12,12, 13,13, 14,14, 15,15, 16,16,
                                 17,17, 18,18, 19,19, 20,20, 21,21, 22,22, 23,23, 24,24,
                                 25,25, 26,26, 27,27, 28,28, 29,29, 30,30, 31,31, 32,32
                                 ),
                       
                       Game2 = c(        1,1, 2,2, 3,3, 4,4, 4,4, 3,3, 2,2, 1,1,
                                         5,5, 6,6, 7,7, 8,8, 8,8, 7,7, 6,6, 5,5,
                                   9,9, 10,10, 11,11, 12,12, 12,12, 11,11, 10,10, 9,9,
                                 13,13, 14,14, 15,15, 16,16, 16,16, 15,15, 14,14, 13,13
                                 ), # Up to 16
  
                       Game3 = c(1,1, 2,2, 2,2, 1,1, 1,1, 2,2, 2,2, 1,1,
                                 3,3, 4,4, 4,4, 3,3, 3,3, 4,4, 4,4, 3,3,
                                 5,5, 6,6, 6,6, 5,5, 5,5, 6,6, 6,6, 5,5,
                                 7,7, 8,8, 8,8, 7,7, 7,7, 8,8, 8,8, 7,7
                                 ), # Start with 8 1s up until 8
  
                       Game4 = c(rep(1,16), rep(2,16), rep(3,16), rep(4,16)), # 16 1s up until 4
                       
### This needs reviewed, but I am making the assumption the bracket quadrants are labeled clock wise
### i.e. W = South, X = East, Y = West, Z = Midwest. This Means in Final 4 you have W vs. Z and X vs. Y.
  
                       Game5 = c(rep(1,16),rep(1,16),rep(2,16),rep(2,16)), # final 4 two games 4 numbers
  
                       Game6 = c(rep(1,64)) # championship 1 game 2 numbers 
)
```

Apply Elo rating simulation
```{r}
elo_vals <- rep(NA, 64)
# Add elo
for(i in 1:64){
  elo_vals[i] <- mm_elo[mm_elo$teams == Tourney24$teams[i],]$elo
  #print(i)
}

Tourney24$elo <- elo_vals

tourney_store <- Tourney24
#Tourney24 <- merge(Tourney24, team_seeds24, by = "seed_s")
#Tourney24 <- Tourney24[c(1,3:10)]
Tourney24 <- tourney_store

set.seed(5882300)

round2 <-  rep(NA, 64)


for (i in 1:32){
  teams <-  Tourney24$teams[which(Tourney24$Game1 == i)]
 
  pred <-   elo.prob(Tourney24$elo[which(Tourney24$teams == teams[1])],
                     Tourney24$elo[which(Tourney24$teams == teams[2])])
  
  sim <- runif(1, min = 0, max = 1)
  if (pred < sim){
    round2[which(Tourney24$teams == teams[2])] <-  Tourney24$Game2[which(Tourney24$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    
    new_elo <- elo.calc(wins.A = 0, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed  
  }
  else{
    round2[which(Tourney24$teams == teams[1])] <- Tourney24$Game2[which(Tourney24$teams == teams[1])]
    new_elo <- elo.calc(wins.A = 1, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed  
    
    }
  # Store new Elo ratings for home team
  Tourney24$elo[which(Tourney24$teams == teams[1])] <- new_elo[1,1]
  # Store new Elo ratings for away team
  Tourney24$elo[which(Tourney24$teams == teams[2])] <- new_elo[1,2]
  }


Tourney24$round2 <- round2


round3 <-  rep(NA, 64)


for (i in 1:16){
  teams <-  Tourney24$teams[which(Tourney24$round2 == i)]
  
  pred <-  elo.prob(Tourney24$elo[which(Tourney24$teams == teams[1])],
                    Tourney24$elo[which(Tourney24$teams == teams[2])])
  sim <- runif(1, min = 0, max = 1)
  if (pred < sim){
    round3[which(Tourney24$teams == teams[2])] <- Tourney24$Game3[which(Tourney24$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    new_elo <- elo.calc(wins.A = 0, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
    
  }
  else{
    round3[which(Tourney24$teams == teams[1])] <- Tourney24$Game3[which(Tourney24$teams == teams[1])]
    new_elo <- elo.calc(wins.A = 1, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed  
  }
  # Store new Elo ratings for home team
  Tourney24$elo[which(Tourney24$teams == teams[1])] <- new_elo[1,1]
  # Store new Elo ratings for away team
  Tourney24$elo[which(Tourney24$teams == teams[2])] <- new_elo[1,2]
}


Tourney24$round3 <- round3



round4 <-  rep(NA, 64)


for (i in 1:8){
  teams <-  Tourney24$teams[which(Tourney24$round3 == i)]

  pred <-  elo.prob(Tourney24$elo[which(Tourney24$teams == teams[1])],
                    Tourney24$elo[which(Tourney24$teams == teams[2])])
  sim <- runif(1, min = 0, max = 1)
  if (pred < sim){
    round4[which(Tourney24$teams == teams[2])] <-  Tourney24$Game4[which(Tourney24$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    new_elo <- elo.calc(wins.A = 0, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
    
  }
  else{
    round4[which(Tourney24$teams == teams[1])] <- Tourney24$Game4[which(Tourney24$teams == teams[1])]
    new_elo <- elo.calc(wins.A = 1, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed  
  }
  # Store new Elo ratings for home team
  Tourney24$elo[which(Tourney24$teams == teams[1])] <- new_elo[1,1]
  # Store new Elo ratings for away team
  Tourney24$elo[which(Tourney24$teams == teams[2])] <- new_elo[1,2]
}


Tourney24$round4 <-  round4


round5 <-  rep(NA, 64)


for (i in 1:4){
  teams <-  Tourney24$teams[which(Tourney24$round4 == i)]
 
  pred <-  elo.prob(Tourney24$elo[which(Tourney24$teams == teams[1])],
                    Tourney24$elo[which(Tourney24$teams == teams[2])])
  sim <- runif(1, min = 0, max = 1)
  if (pred < sim){
    round5[which(Tourney24$teams == teams[2])] <-  Tourney24$Game5[which(Tourney24$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    new_elo <- elo.calc(wins.A = 0, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
    
  }
  else{
    round5[which(Tourney24$teams == teams[1])] <- Tourney24$Game5[which(Tourney24$teams == teams[1])]
    new_elo <- elo.calc(wins.A = 1, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
    
  }
  # Store new Elo ratings for home team
  Tourney24$elo[which(Tourney24$teams == teams[1])] <- new_elo[1,1]
  # Store new Elo ratings for away team
  Tourney24$elo[which(Tourney24$teams == teams[2])] <- new_elo[1,2]
}



Tourney24$round5 <- round5



round6 <-  rep(NA, 64) 


for (i in 1:2){
  teams <-  Tourney24$teams[which(Tourney24$round5 == i)]
 
  pred <-  elo.prob(Tourney24$elo[which(Tourney24$teams == teams[1])],
                    Tourney24$elo[which(Tourney24$teams == teams[2])])
  sim <- runif(1, min = 0, max = 1)
  if (pred < sim){
    round6[which(Tourney24$teams == teams[2])] <-  Tourney24$Game6[which(Tourney24$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    new_elo <- elo.calc(wins.A = 0, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
    
  }
  else{
    round6[which(Tourney24$teams == teams[1])] <- Tourney24$Game6[which(Tourney24$teams == teams[1])]
    new_elo <- elo.calc(wins.A = 1, # Select game outcome
                        elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
  }
  # Store new Elo ratings for home team
  Tourney24$elo[which(Tourney24$teams == teams[1])] <- new_elo[1,1]
  # Store new Elo ratings for away team
  Tourney24$elo[which(Tourney24$teams == teams[2])] <- new_elo[1,2]
}



Tourney24$round6 <- round6

# Predict Championship
Champ <-  rep(NA, 64)


  teams <-  Tourney24$teams[which(Tourney24$round6 == 1)]
  pred <-  elo.prob(Tourney24$elo[which(Tourney24$teams == teams[1])],
                  Tourney24$elo[which(Tourney24$teams == teams[2])])
  sim <- runif(1, min = 0, max = 1)
  if (pred < sim){
  Champ[which(Tourney24$teams == teams[2])] <-  Tourney24$Game6[which(Tourney24$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
  new_elo <- elo.calc(wins.A = 0, # Select game outcome
                      elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                      elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                      k= 20) # Set update speed 
  
}else{
  Champ[which(Tourney24$teams == teams[1])] <- Tourney24$Game6[which(Tourney24$teams == teams[1])]
  new_elo <- elo.calc(wins.A = 1, # Select game outcome
                      elo.A = Tourney24$elo[which(Tourney24$teams == teams[1])], # Set Elo for team 1
                      elo.B = Tourney24$elo[which(Tourney24$teams == teams[2])], # Set Elo for team 2
                      k= 20) # Set update speed 
}
# Store new Elo ratings for home team
Tourney24$elo[which(Tourney24$teams == teams[1])] <- new_elo[1,1]
# Store new Elo ratings for away team
Tourney24$elo[which(Tourney24$teams == teams[2])] <- new_elo[1,2]
Tourney24$Champ <- Champ
print(Tourney24$teams[which(Tourney24$Champ == 1)])
```

### Multiple Simulations

Let's create the tournament system
```{r}
Tourney24 <-  data.frame(teams = c(#East:
  "UConn", # 1
  "Stetson", # 16
  "Iowa State", # 2
  "South Dakota State", # 15
  "Illinois", # 3 
  "Morehead State", # 14
  "Auburn", # 4
  "Yale", # 13
  "San Diego State", # 5
  "UAB", # 12
  "BYU", # 6
  "Duquesne", # 11
  "Washington State", # 7 
  "Drake", # 10
  "Florida Atlantic", # 8 
  "Northwestern", # 9  
  
  #West:
  "North Carolina", # 1
  "Wagner", # 16 
  "Arizona", # 2
  "Long Beach State", # 15
  "Baylor", # 3
  "Colgate", # 14
  "Alabama", # 4
  "Charleston", # 13
  "Saint Mary's", # 5
  "Grand Canyon", # 12
  "Clemson", # 6
  "New Mexico", # 11 
  "Dayton", # 7
  "Nevada", # 10
  "Mississippi State", # 8
  "Michigan State", # 9
  
  #South:
  "Houston", # 1
  "Longwood", # 16
  "Marquette", # 2
  "Western Kentucky", # 15
  "Kentucky", # 3
  "Oakland", # 14
  "Duke", # 4
  "Vermont", # 13
  "Wisconsin", # 5
  "James Madison", # 12
  "Texas Tech", # 6
  "NC State", # 11
  "Florida", # 7
  "Colorado", # 10, or Boise State
  "Nebraska", # 8
  "Texas A&M", # 9
  
  #Midwest:
  "Purdue", # 1
  "Grambling", # 16
  "Tennessee", # 2
  "Saint Peter's", # 15
  "Creighton", # 3
  "Akron", # 14
  "Kansas", # 4
  "Samford", # 13
  "Gonzaga", # 5
  "McNeese", # 12
  "South Carolina", # 6
  "Oregon", # 11
  "Texas", # 7
  "Colorado State", # 10
  "Utah State", # 8
  "TCU" # 9
  ), 
  
                       seed_s = c("W01", "W16", "W02", "W15", "W03", "W14", "W04", "W13",
                                  "W05", "W12", "W06", "W11", "W07", "W10", "W08", "W09",
                                  "X01", "X16", "X02", "X15", "X03", "X14", "X04", "X13",
                                  "X05", "X12", "X06", "X11", "X07", "X10", "X08", "X09",
                                  "Y01", "Y16", "Y02", "Y15", "Y03", "Y14", "Y04", "Y13", 
                                  "Y05", "Y12", "Y06", "Y11", "Y07", "Y10", "Y08", "Y09",
                                  "Z01", "Z16", "Z02", "Z15", "Z03", "Z14", "Z04", "Z13",
                                  "Z05", "Z12", "Z06", "Z11", "Z07", "Z10", "Z08", "Z09"
                                  ), 
                       
                       
                       # Work from outsides of range building into the median to create WO1 vs. WO16 and so on reference the march madness bracket. 
                       
                       Game1 = c( 1,1,   2,2,   3,3,   4,4,   5,5,   6,6,   7,7,   8,8,
                                  9,9,  10,10, 11,11, 12,12, 13,13, 14,14, 15,15, 16,16,
                                 17,17, 18,18, 19,19, 20,20, 21,21, 22,22, 23,23, 24,24,
                                 25,25, 26,26, 27,27, 28,28, 29,29, 30,30, 31,31, 32,32
                                 ),
                       
                       Game2 = c(        1,1, 2,2, 3,3, 4,4, 4,4, 3,3, 2,2, 1,1,
                                         5,5, 6,6, 7,7, 8,8, 8,8, 7,7, 6,6, 5,5,
                                   9,9, 10,10, 11,11, 12,12, 12,12, 11,11, 10,10, 9,9,
                                 13,13, 14,14, 15,15, 16,16, 16,16, 15,15, 14,14, 13,13
                                 ), # Up to 16
  
                       Game3 = c(1,1, 2,2, 2,2, 1,1, 1,1, 2,2, 2,2, 1,1,
                                 3,3, 4,4, 4,4, 3,3, 3,3, 4,4, 4,4, 3,3,
                                 5,5, 6,6, 6,6, 5,5, 5,5, 6,6, 6,6, 5,5,
                                 7,7, 8,8, 8,8, 7,7, 7,7, 8,8, 8,8, 7,7
                                 ), # Start with 8 1s up until 8
  
                       Game4 = c(rep(1,16), rep(2,16), rep(3,16), rep(4,16)), # 16 1s up until 4
                       
### This needs reviewed, but I am making the assumption the bracket quadrants are labeled clock wise
### i.e. W = South, X = East, Y = West, Z = Midwest. This Means in Final 4 you have W vs. Z and X vs. Y.
  
                       Game5 = c(rep(1,16),rep(1,16),rep(2,16),rep(2,16)), # final 4 two games 4 numbers
  
                       Game6 = c(rep(1,64)) # championship 1 game 2 numbers 
)

#Apply Elo rating simulation
elo_vals <- rep(NA, 64)
# Add elo
for(i in 1:64){
  elo_vals[i] <- mm_elo[mm_elo$teams == Tourney24$teams[i],]$elo
  #print(i)
}

Tourney24$elo <- elo_vals
```

Let's create a tournament simulation function
```{r}
march_madness_sim <- function(team_db, k){

  Tourney <- team_db
  

  round2 <-  rep(NA, 64)


  for (i in 1:32){
    teams <-  Tourney$teams[which(Tourney$Game1 == i)]
 
    pred <-   elo.prob(Tourney$elo[which(Tourney$teams == teams[1])],
                       Tourney$elo[which(Tourney$teams == teams[2])])
  
    sim <- runif(1, min = 0, max = 1)
    if (pred < sim){
      round2[which(Tourney$teams == teams[2])] <-  Tourney$Game2[which(Tourney$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    
      new_elo <- elo.calc(wins.A = 0, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed  
    }
    else{
      round2[which(Tourney$teams == teams[1])] <- Tourney$Game2[which(Tourney$teams == teams[1])]
      new_elo <- elo.calc(wins.A = 1, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed  
    
      }
    # Store new Elo ratings for home team
    Tourney$elo[which(Tourney$teams == teams[1])] <- new_elo[1,1]
    # Store new Elo ratings for away team
    Tourney$elo[which(Tourney$teams == teams[2])] <- new_elo[1,2]
    }


  Tourney$round2 <- round2


  round3 <-  rep(NA, 64)


  for (i in 1:16){
    teams <-  Tourney$teams[which(Tourney$round2 == i)]
  
    pred <-  elo.prob(Tourney$elo[which(Tourney$teams == teams[1])],
                      Tourney$elo[which(Tourney$teams == teams[2])])
    sim <- runif(1, min = 0, max = 1)
    if (pred < sim){
      round3[which(Tourney$teams == teams[2])] <- Tourney$Game3[which(Tourney$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
      new_elo <- elo.calc(wins.A = 0, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed 
    
    }
    else{
      round3[which(Tourney$teams == teams[1])] <- Tourney$Game3[which(Tourney$teams == teams[1])]
      new_elo <- elo.calc(wins.A = 1, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed  
    }
    # Store new Elo ratings for home team
    Tourney$elo[which(Tourney$teams == teams[1])] <- new_elo[1,1]
    # Store new Elo ratings for away team
    Tourney$elo[which(Tourney$teams == teams[2])] <- new_elo[1,2]
  }


  Tourney$round3 <- round3



  round4 <-  rep(NA, 64)


  for (i in 1:8){
    teams <-  Tourney$teams[which(Tourney$round3 == i)]

    pred <-  elo.prob(Tourney$elo[which(Tourney$teams == teams[1])],
                      Tourney$elo[which(Tourney$teams == teams[2])])
    sim <- runif(1, min = 0, max = 1)
    if (pred < sim){
      round4[which(Tourney$teams == teams[2])] <-  Tourney$Game4[which(Tourney$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
      new_elo <- elo.calc(wins.A = 0, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed 
    
    }
    else{
      round4[which(Tourney$teams == teams[1])] <- Tourney$Game4[which(Tourney$teams == teams[1])]
      new_elo <- elo.calc(wins.A = 1, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed  
    }
    # Store new Elo ratings for home team
    Tourney$elo[which(Tourney$teams == teams[1])] <- new_elo[1,1]
    # Store new Elo ratings for away team
    Tourney$elo[which(Tourney$teams == teams[2])] <- new_elo[1,2]
  }


  Tourney$round4 <-  round4


  round5 <-  rep(NA, 64)


  for (i in 1:4){
    teams <-  Tourney$teams[which(Tourney$round4 == i)]
 
    pred <-  elo.prob(Tourney$elo[which(Tourney$teams == teams[1])],
                      Tourney$elo[which(Tourney$teams == teams[2])])
    sim <- runif(1, min = 0, max = 1)
    if (pred < sim){
      round5[which(Tourney$teams == teams[2])] <-  Tourney$Game5[which(Tourney$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
      new_elo <- elo.calc(wins.A = 0, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed 
    
    }
    else{
      round5[which(Tourney$teams == teams[1])] <- Tourney$Game5[which(Tourney$teams == teams[1])]
      new_elo <- elo.calc(wins.A = 1, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed 
    
    }
    # Store new Elo ratings for home team
    Tourney$elo[which(Tourney$teams == teams[1])] <- new_elo[1,1]
    # Store new Elo ratings for away team
    Tourney$elo[which(Tourney$teams == teams[2])] <- new_elo[1,2]
  }



  Tourney$round5 <- round5



  round6 <-  rep(NA, 64) 


  for (i in 1:2){
    teams <-  Tourney$teams[which(Tourney$round5 == i)]
 
    pred <-  elo.prob(Tourney$elo[which(Tourney$teams == teams[1])],
                    Tourney$elo[which(Tourney$teams == teams[2])])
    sim <- runif(1, min = 0, max = 1)
    if (pred < sim){
      round6[which(Tourney$teams == teams[2])] <-  Tourney$Game6[which(Tourney$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
      new_elo <- elo.calc(wins.A = 0, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed 
    
    }
    else{
      round6[which(Tourney$teams == teams[1])] <- Tourney$Game6[which(Tourney$teams == teams[1])]
      new_elo <- elo.calc(wins.A = 1, # Select game outcome
                          elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                          elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                          k= 20) # Set update speed 
    }
    # Store new Elo ratings for home team
    Tourney$elo[which(Tourney$teams == teams[1])] <- new_elo[1,1]
    # Store new Elo ratings for away team
    Tourney$elo[which(Tourney$teams == teams[2])] <- new_elo[1,2]
  }



  Tourney$round6 <- round6

  # Predict Championship
  Champ <-  rep(NA, 64)


    teams <-  Tourney$teams[which(Tourney$round6 == 1)]
    
    # Join data for two teams in game to single row:
    
    
    
    pred <-  elo.prob(Tourney$elo[which(Tourney$teams == teams[1])],
                    Tourney$elo[which(Tourney$teams == teams[2])])
    sim <- runif(1, min = 0, max = 1)
    if (pred < sim){
    Champ[which(Tourney$teams == teams[2])] <-  Tourney$Game6[which(Tourney$teams == teams[2])] # assign round 2 game to team 2 if they were to win the game
    new_elo <- elo.calc(wins.A = 0, # Select game outcome
                        elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
  
  }
  else{
    Champ[which(Tourney$teams == teams[1])] <- Tourney$Game6[which(Tourney$teams == teams[1])]
    new_elo <- elo.calc(wins.A = 1, # Select game outcome
                        elo.A = Tourney$elo[which(Tourney$teams == teams[1])], # Set Elo for team 1
                        elo.B = Tourney$elo[which(Tourney$teams == teams[2])], # Set Elo for team 2
                        k= 20) # Set update speed 
  }
  # Store new Elo ratings for home team
  Tourney$elo[which(Tourney$teams == teams[1])] <- new_elo[1,1]
  # Store new Elo ratings for away team
  Tourney$elo[which(Tourney$teams == teams[2])] <- new_elo[1,2]
  Tourney$Champ <- Champ

  return(Tourney)
}
```


Simulating the tournament 10,000 times
```{r}
# Create a vector to store tournament winner
winner <- rep(NA, 10000)

round2 <- round3 <- round4 <- round5 <- round6 <- Champ <- rep(0, nrow(Tourney24))

bracket_res <- data.frame(Tourney24$teams, Tourney24$seed_s)
# For 1 to 10,000
for(x in 1:10000){
  # Set seed for reproducability
  set.seed(x + 5882300)
  # Simulate tournament
  temp <- march_madness_sim(team_db = Tourney24, k = 50)
  # Store sweet 16 progression
  round2 <- round2 + !is.na(temp$round2)
  # Store sweet 16 progression
  round3 <- round3 + !is.na(temp$round3)
  # Store elite 8 progression
  round4 <- round4 + !is.na(temp$round4)
  # Store final 4 progression
  round5 <- round5 + !is.na(temp$round5)
  # Store championship progression
  round6 <- round6 + !is.na(temp$round6)
  # Store victory
  Champ <- Champ + !is.na(temp$Champ)
  
  # Extract tournament winner
  winner[x] <- temp$teams[which(temp$Champ == 1)]
}

#Store overall bracket results
bracket_res$round32 <- round2
bracket_res$sweet16 <- round3
bracket_res$elite8 <- round4
bracket_res$final4 <- round5
bracket_res$championship <- round6
bracket_res$winner <- Champ
  
names(bracket_res)[c(1,2)] <- c("teams", "seed")
```

Let's compute the probabilities that each team will make each round
```{r}
bracket_prob <- bracket_res[,c(3:8)]/1000
```


```{r}
bracket_temp <- bracket_prob
bracket_temp[,] <- 0

for(i in 1:max(Tourney24$Game1)){
  bracket_temp$round32[which(Tourney24$Game1 == i)][which.max(bracket_prob$round32[which(Tourney24$Game1 == i)])] <- 1
}

for(i in 1:max(Tourney24$Game2)){
  bracket_temp$sweet16[which(Tourney24$Game2 == i)][which.max(bracket_prob$sweet16[which(Tourney24$Game2 == i)])] <- 1
}

for(i in 1:max(Tourney24$Game3)){
  bracket_temp$elite8[which(Tourney24$Game3 == i)][which.max(bracket_prob$elite8[which(Tourney24$Game3 == i)])] <- 1
}

for(i in 1:max(Tourney24$Game4)){
  bracket_temp$final4[which(Tourney24$Game4 == i)][which.max(bracket_prob$final4[which(Tourney24$Game4 == i)])] <- 1
}

for(i in 1:max(Tourney24$Game5)){
  bracket_temp$championship[which(Tourney24$Game5 == i)][which.max(bracket_prob$championship[which(Tourney24$Game5 == i)])] <- 1
}

for(i in 1:max(Tourney24$Game6)){
  bracket_temp$winner[which(Tourney24$Game6 == i)][which.max(bracket_prob$winner[which(Tourney24$Game6 == i)])] <- 1
}
```

```{r}
bracket_final <- cbind.data.frame(bracket_res, bracket_temp)
bracket_final
```

Save file as a .csv
```{r}
write.csv(bracket_final, "bracket_elo_prediction.csv")
```

### XGBOOST SIM

```{r}
knight <- read.csv("TournamentTeamHistoryMetrics.csv")
```

```{r}
knight <- knight %>%
  filter(YEAR == 2024) %>%
  select(-c(3:5,23)) 
summary(knight)

write.csv(knight, "test.csv", row.names = FALSE)
```



```{r}
prepare_features <- function(team1, team2, Tourney) {
    team1_data <- Tourney[which(Tourney$teams == team1), ]
    team2_data <- Tourney[which(Tourney$teams == team2), ]

    data.frame(
        elo_diff = team1_data$elo - team2_data$elo,
        seed_diff = team1_data$seed - team2_data$seed,
        win_pct_diff = team1_data$win_pct - team2_data$win_pct,
        off_rating_diff = team1_data$off_rating - team2_data$off_rating,
        def_rating_diff = team1_data$def_rating - team2_data$def_rating,
        avg_points_scored_diff = team1_data$avg_points_scored - team2_data$avg_points_scored,
        avg_points_allowed_diff = team1_data$avg_points_allowed - team2_data$avg_points_allowed,
        experience_diff = team1_data$tournament_experience - team2_data$tournament_experience,
        
        # Add other features as needed
    )
}

march_madness_sim2 <- function(team_db, bst_final) {
    Tourney <- team_db
    round2 <- rep(NA, nrow(Tourney))

    for (i in 1:32) {
        teams <- Tourney$teams[which(Tourney$Game1 == i)]

        # Prepare the features for the XGBoost model
        features <- prepare_features(teams[1], teams[2], Tourney)
        
        # convert to xbg.dmatrix
        

        
        # Get the prediction from the XGBoost model
        pred <- predict(bst_final, as.matrix(features), type = "prob")

        sim <- runif(1)
        if (pred < sim) {
            round2[which(Tourney$teams == teams[2])] <- Tourney$Game2[which(Tourney$teams == teams[2])]
        } else {
            round2[which(Tourney$teams == teams[1])] <- Tourney$Game2[which(Tourney$teams == teams[1])]
        }
    }

    Tourney$round2 <- round2
    
    round3 <- rep(NA, nrow(Tourney))

    for (i in 1:16) {
        teams <- Tourney$teams[which(Tourney$round2 == i)]
  
    # Check if there are two teams to compare, to avoid errors
       if (length(teams) == 2) {
        # Prepare the features for the XGBoost model
            features <- prepare_features(teams[1], teams[2], Tourney)
  
        # Get the prediction from the XGBoost model
         pred <- predict(bst_final, as.matrix(features), type = "prob")
  
         sim <- runif(1)
        if (pred < sim) {
            round3[which(Tourney$teams == teams[2])] <- Tourney$Game3[which(Tourney$teams == teams[2])]
        } else {
            round3[which(Tourney$teams == teams[1])] <- Tourney$Game3[which(Tourney$teams == teams[1])]
        }
      }
   }

   Tourney$round3 <- round3
   
   
    round4 <- rep(NA, nrow(Tourney))

    for (i in 1:8) {
    teams <- Tourney$teams[which(Tourney$round3 == i)]

    # Ensure we have two teams for the comparison
        if (length(teams) == 2) {
        # Prepare the features for the XGBoost model
        features <- prepare_features(teams[1], teams[2], Tourney)
    
        # Get the prediction from the XGBoost model
        pred <- predict(bst_final, as.matrix(features), type = "prob")
  
        sim <- runif(1)
        if (pred < sim) {
            round4[which(Tourney$teams == teams[2])] <- Tourney$Game4[which(Tourney$teams == teams[2])]
        } else {
            round4[which(Tourney$teams == teams[1])] <- Tourney$Game4[which(Tourney$teams == teams[1])]
        }
       }
    }

    Tourney$round4 <- round4
    
    round5 <- rep(NA, nrow(Tourney))

    for (i in 1:4) {
        teams <- Tourney$teams[which(Tourney$round4 == i)]

        if (length(teams) == 2) {
        # Prepare the features for the XGBoost model
        features <- prepare_features(teams[1], teams[2], Tourney)

        # Get the prediction from the XGBoost model
        pred <- predict(bst_final, as.matrix(features), type = "prob")

        sim <- runif(1)
        if (pred < sim) {
            round5[which(Tourney$teams == teams[2])] <- Tourney$Game5[which(Tourney$teams == teams[2])]
        } else {
            round5[which(Tourney$teams == teams[1])] <- Tourney$Game5[which(Tourney$teams == teams[1])]
        }
      }
    }

    Tourney$round5 <- round5
    
    round6 <- rep(NA, nrow(Tourney))

    for (i in 1:2) {
      teams <- Tourney$teams[which(Tourney$round5 == i)]

        if (length(teams) == 2) {
        # Prepare the features for the XGBoost model
        features <- prepare_features(teams[1], teams[2], Tourney)

        # Get the prediction from the XGBoost model
        pred <- predict(bst_final, as.matrix(features), type = "prob")

        sim <- runif(1)
        if (pred < sim) {
            round6[which(Tourney$teams == teams[2])] <- Tourney$Game6[which(Tourney$teams == teams[2])]
        } else {
            round6[which(Tourney$teams == teams[1])] <- Tourney$Game6[which(Tourney$teams == teams[1])]
        }
      }
    }

    Tourney$round6 <- round6

    Champ <- rep(NA, nrow(Tourney))

    teams <- Tourney$teams[which(Tourney$round6 == 1)]

if (length(teams) == 2) {
    # Prepare the features for the XGBoost model
    features <- prepare_features(teams[1], teams[2], Tourney)

    # Get the prediction from the XGBoost model
    pred <- predict(bst_final, as.matrix(features), type = "prob")

    sim <- runif(1)
    if (pred < sim) {
        Champ[which(Tourney$teams == teams[2])] <- Tourney$Game6[which(Tourney$teams == teams[2])]
    } else {
        Champ[which(Tourney$teams == teams[1])] <- Tourney$Game6[which(Tourney$teams == teams[1])]
      }
    }

    Tourney$Champ <- Champ






    # Repeat similar logic for round3 to round6...
    # Ensure you update the game numbers and rounds appropriately

  return(Tourney)
}


```

```{r}
# Ensure bst_final model is loaded and available

# Create a vector to store tournament winner
winner <- rep(NA, 10000)

# Initialize counters for each round's progression
round2 <- round3 <- round4 <- round5 <- round6 <- Champ <- rep(0, nrow(Tourney24))

# Data frame to store the results
bracket_res <- data.frame(Tourney24$teams, Tourney24$seed_s)
merged_data <- merge(Tourney24, knight, by.x = "teams", by.y = "TEAM", all.x = TRUE)

# Simulate the tournament 10,000 times
for(x in 1:10000) {
    set.seed(x + 5882300)  # for reproducibility
    temp <- march_madness_sim(team_db = merged_data, bst_final)

    # Accumulate the number of times each team progresses in each round
    round2 <- round2 + !is.na(temp$round2)
    round3 <- round3 + !is.na(temp$round3)
    round4 <- round4 + !is.na(temp$round4)
    round5 <- round5 + !is.na(temp$round5)
    round6 <- round6 + !is.na(temp$round6)
    Champ <- Champ + !is.na(temp$Champ)
  
    # Determine the tournament winner for each simulation
    winner[x] <- temp$teams[which(temp$Champ == 1)]
}

# Store overall bracket results
bracket_res$round32 <- round2
bracket_res$sweet16 <- round3
bracket_res$elite8 <- round4
bracket_res$final4 <- round5
bracket_res$championship <- round6
bracket_res$winner <- Champ

names(bracket_res)[c(1,2)] <- c("teams", "seed")

```

